The Evolution of Phenotypically Invariant Genetic Networks

### 2021-08-11 kalman_walk_paper.tex
bib fixes

In this paper we use results on mathematical nonidentifiability from \llabel{r21}
linear systems theory to study how gene regulatory networks can be modified
while retaining the same function, and the possible implications for speciation.
% Specifically, we would like to understand the degree of which gene regulatory network architectures are functionally equivalent and how this relates to network size and efficiency.
If system architectures are not functionally unique,
can this open up neutral evolutionary paths,
and do species explore these paths
through the process termed \emph{developmental system drift} \citep[by][]{true2001developmental}?
Is this fast enough to contribute meaningfully to speciation?
To do this, we describe results on linear dynamical systems which give
an analytical description of the set of 
all linear gene network architectures that yield identical phenotypes,
and use quantitative genetics theory to estimate the speed at which system drift can lead to reproductive incompatibility and hence speciation.
%
In this model,
a population diffuses along the neutral ridges of a high-dimensional space of possible system parameters,
in a similar vein as \emph{holey landscape} models \citep{gavrilets1997evolution, yamaguchi2013first, pina2019does}.

@@

%\section*{Acknowledgements}
%
%    We would like to thank Sergey Nuzhdin, Stevan Arnold, Michael Turelli, Patrick Phillips, Erik Lundgren and Hossein Asgharian for valuable discussion. 
%    We would also like to thank Nick Barton, Sarah Signor, Todd Parsons, and Joachim Hermisson for very helpful comments on the manuscript.
%    Work on this project was supported by funds from
%    the Sloan Foundation and the NSF (under DBI-1262645) to PR.

>>>>

In this paper we use results on mathematical nonidentifiability from \llabel{r21}
linear systems theory to study how gene regulatory networks can be modified
while retaining the same function, and the possible implications for speciation.
% Specifically, we would like to understand the degree of which gene regulatory network architectures are functionally equivalent and how this relates to network size and efficiency.
If system architectures are not functionally unique,
can this open up neutral evolutionary paths,
and do species explore these paths
through the process termed \emph{developmental system drift} \citep[by][]{true2001developmental}?
Is this fast enough to contribute meaningfully to speciation?
To do this, we describe results on linear dynamical systems which give
an analytical description of the set of 
all linear gene network architectures that yield identical phenotypes,
and use quantitative genetics theory to estimate the speed at which system drift can lead to reproductive incompatibility and hence speciation.
%
In this model,
a population diffuses along the neutral ridges of a high-dimensional space of possible system parameters,
in a similar vein as \emph{holey landscape} models \citep{gavrilets1997evolution, yamaguchi2013first}.

@@

\section*{Acknowledgements}

We would like to thank Sergey Nuzhdin, Stevan Arnold, Michael Turelli, Patrick Phillips, Erik Lundgren and Hossein Asgharian for valuable discussion. 
We would also like to thank Nick Barton, Sarah Signor, Todd Parsons, and Joachim Hermisson for very helpful comments on the manuscript.
Work on this project was supported by funds from
the Sloan Foundation and the NSF (under DBI-1262645) to PR.
